---
title: You Know Nothing Jon Snow
author: David Earl
date: '2017-10-01'
slug: you-know-nothing-jon-snow
categories: []
tags:
  - network-graph
weight: 3
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE, warning=FALSE, message=FALSE)
```

**Intro**

The world created by George R. R. Martin in the Song of Ice and Fire series is 
full of warring factions and complex relationships between characters. This 
analysis attempts to build a network graph of the world from the raw text of 
the books with no foreknowledge of the plot or characters.  

**Getting Started**

Each book was exported from an ebook reader as a plain text file. The table of 
contents, introduction, and appendixes were removed and extra white space was 
trimmed. The books were then read in and collapsed into a single character 
string. 

```{r}
load_books <- function(text_dir = '.', encoding = 'ANSI') {
  books <- c()
  for (text in list.files(text_dir)) {
    books[[text]] <- readLines(paste0(text_dir, text), encoding = encoding)
  }
  return(books)
}

collapse_books <- function(books) {
  for (book in seq(length(books))) {
    books[[book]] <- paste(books[[book]], sep = '', collapse = '')
  }
  return(paste(books, collapse = ''))
}

books <- load_books('./texts/')
books <- collapse_books(books)
```

Next, taking advantage of the structure of dialog I first tried identifying 
characters by finding capitalized words that appear after end quotes `”` 
using the regular expressions `'” [A-Z][a-z]+'`.

After some testing on some sample text this approach looks promising. However, 
it looks like full names are being truncated.

`[1] "Gared"  "Ser"  "Royce"  "Gared"  "Royce"  "Gared"`

Expanding the number of time the pattern is matched returns titles and full 
names. Also one of the people groups in the books use 'kh' to indicate proper 
nouns so the pattern will be adjusted to account for those cases as 
well. `'” (([A-Z]|kh)[a-z]+ ){1,3}'`

`[1] "Gared"  "Ser Waymar Royce"  "Royce" "Gared" "Royce"`          

While the `stringr` package is awesome and typically more convenient, I've found 
the base R string functions are more forgiving when dealing with text sources 
not encoded in UTF-8.  The function below uses `gregexpr` and `substr` to 
extract strings matching the regex pattern and returns a character vector. 
It also stores the position of the match using the names attribute of the 
vector and can filter out stopwords to simplify later analysis.

```{r}
extract_pattern <- function(string,
                            pattern = "” (([A-Z]|kh)[a-z]+ ){1,3}",
                            offset_start = 2,
                            offset_end = 2,
                            stopwords = NULL) {
  matches <- gregexpr(pattern, string)
  match_lengths <- attributes(matches[[1]])$match.length
  matches <- unlist(matches)
  extracted <- c()
  for (i in seq(length(matches))) {
    match_str <- substr(x = string, 
                        start = matches[i] + offset_start,
                        stop = matches[i] + match_lengths[[i]] - offset_end)
    if (!tolower(match_str) %in% stopwords) {
       extracted[[paste(matches[i])]] <-  match_str
     }
  }
  return(extracted)
}

extra_stopwords <- c('none', 'one', 'sometimes', 'must',
                     'still', 'another', 'better', 'yet',
                     'together', 'though', 'perhaps', 'unless', 'maybe',
                     'worse', 'whenever', 'even', 'oh', '')

stopwords <- c(tm::stopwords(), extra_stopwords)

speakers <- extract_pattern(books, stopwords = stopwords)
head(sort(table(speakers), decreasing = T), 50)
```

Spot checking with `table` higlights a few issues. First proper titles and last 
names are causing there to be multiple entries for a single character so I will 
want to combine them by using standard names. 

```{r}
titles <- c('Captain', 'Grand', 'Khal', 'King', 'Lady','Lord',
            'Maester', 'Magister', 'Prince', 'Princess', 'Queen',
            'Septa', 'Ser')

swap_to <- c('Jon', 'Sam', 'Dany', 'Ned', 'Mance Rayder',
             'Meera', 'Mirri Maz Duur', 'Lysa')

swap_from <- c('Jon Snow', 'Samwell Tarly', 'Daenerys', 'Eddard Stark', 
               'Mance', 'Meera Reed', 'Mirri', 'Lysa Arryn')

clean_speakers <- function(speakers, titles, swap_to, swap_from) {
  speakers <- gsub(paste(titles, collapse = '|'), "", speakers)
  speakers <- trimws(speakers)
  for (i in seq(length(swap_to))) {
    speakers[which(speakers == swap_from[i])] <- swap_to[i]
  }
  return(speakers)
}

cl_speakers <- clean_speakers(speakers, titles = titles, 
                           swap_to = swap_to, swap_from = swap_from)

head(sort(table(cl_speakers), decreasing = T), 50)
```

Next the parsed dialogue can be used to build simple network representations. 
Each character who appears as speaker will be a node and edges will be defined 
for each character they spoke before and weighted by the number of times they 
spoke before the other character.  I also want to filter out minor characters 
and interactions.  I will be using the `igraph` package which can build a 
network object from a data.frame of nodes and other attributes.

```{r}
node_df <- function(speakers, min_edges = 3, min_edge_weight = 2) {
  people <- unique(speakers)
  spoke_before <- list()
  for (person in people) {
    if (length(which(speakers == person)) > min_edges) {
      tbl <- table(speakers[which(speakers == person)+1])
      df <- as.data.frame(tbl[tbl > min_edge_weight])
      if (nrow(df) > min_edges) {
        df$person <- person
        df <- df[c(3,1,2)]
        names(df) <- c('node1', 'node2', 'weight')
        spoke_before[[person]] <- df
      }
    }
  }
  return(dplyr::bind_rows(spoke_before))
}

build_network_graph <- function(node_df, dim = 3, niter = 2500) {
  net <- igraph::graph_from_data_frame(node_df)
  net <- igraph::simplify(net, remove.multiple = F)
  lay <- igraph::layout_with_fr(net, dim = dim, niter = niter)
  return(list('network' = net, 'layout' = lay))
}

speaker_nodes <- node_df(cl_speakers)
network <- build_network_graph(speaker_nodes)
```

```{r}
net <- network[['network']]

lay <- network[['layout']]

char_names <- igraph::vertex_attr(net)$name

node_size <- sapply(igraph::V(net)$name, function(x) {
  length(igraph::E(net)[from(x)]) + 1
}) 

edge_weights <- igraph::edge_attr(net)$weight

norm_edge_weights <- round(edge_weights/max(edge_weights), 1)

edges <- igraph::get.edges(net, igraph::E(net))

xs <- c()
ys <- c()
zs <- c()
for (i in seq(nrow(edges))) {
  xs <- c(xs, lay[edges[i,1],1], lay[edges[i,2],1], NA)
  ys <- c(ys, lay[edges[i,1],2], lay[edges[i,2],2], NA)
  zs <- c(zs, lay[edges[i,1],3], lay[edges[i,2],3], NA)
}

rep_weights <- unlist(lapply(norm_edge_weights, rep, 3))

blues <- which(rep_weights %in% c(0.3, 0.4, 0.5))
purples <- which(rep_weights %in% c(0.6, 0.7, 0.8))
reds <- which(rep_weights %in% c(0.9, 1))
not_greys <- c(blues, purples, reds)
```

```{r}
library(plotly)
plot_ly(x = xs[-not_greys], 
        y = ys[-not_greys], 
        z = zs[-not_greys], 
        type = 'scatter3d',
        hoverinfo = 'none',
        mode = 'lines', 
        line = list(color = 'grey',
                    width = 1)
        ) %>%
  add_trace(x = xs[blues],
            y = ys[blues],
            z = zs[blues], 
            mode = 'lines', 
            line = list(color = 'blue', 
                         width = 3)
            ) %>%
  add_trace(x = xs[purples], 
            y = ys[purples], 
            z = zs[purples], 
            mode = 'lines', 
            line = list(color = 'green',
                         width = 4)
            ) %>%
  add_trace(x = xs[reds], 
            y = ys[reds],
            z = zs[reds], 
            mode = 'lines', 
            line = list(color = 'red', 
                         width = 5)
            ) %>%
  add_trace(x = lay[,1], 
            y = lay[,2], 
            z = lay[,3],
            mode = 'lines+markers',
            text = char_names,
            textposition = 'topleft',
            #textfont = list(family = 'arial', size = 10),
            hoverinfo = 'text',
            marker = list(size = node_size, 
                          color = 'orange', 
                          sizemin = 10))
```


